taxadb: A High-Performance Local Taxonomic Database Interface

### 2020-08-04 paper/paper.Rmd
WIP

@@

The workhorse function of `neonstore` is `neon_download()`, which downloads all files identified by a given product code, optionally constrained to specific site, date range, or file name pattern.  For instance, we can download all of the NEON landbird survey data, identified as NEON data product `DP1.10003.001`:

@@

`neon_download()`

The three-digit code identifies this as a Level 1 data product (DP1), product number 10003, revision 001, indicating this is the first and so far the only protocol used for bird survey sampling.

Now, view your store of NEON products:

```{r }
neon_store()
```

These will persist between sessions, so you only need to download once
or to retrieve updates. `neon_store()` can take arguments to filter by product
or pattern in table name, e.g. `neon_store(table = "brd")`.

Once you determine the table of interest, you can
read in all the component tables into a single `data.frame`

>>>>

### A local store for neon files

The location of the local store can be configured depending on the context.  The default behavior is to rely on a user-specific location appropriate for the operating system, e.g. `~/.local/share/neonstore` on Linux, `C:\Users\<username>\AppData\Local\neonstore` on Windows, or `~/Library/Application Support/neonstore` on Mac OS X, as determined by the `rappdirs` R package which provides appropriate modern defaults [@rappdirs].  This approach works well for a single user setup such as a laptop, where a user can share the store across multiple projects in different directories.  On a server system that is accessed by multiple users, an lower-level location that permits all users to read and write files to the store may be preferable.  Because `neonstore` only writes raw NEON data to the store, there is no risk of overwriting work of other users, while the shared storage can reduce consumption of redundant download and storage.  In yet other cases, a user may prefer a path that is specific to a single project.

It is of course possible to use a temporary directory such as `tempdir()` for the storage location.  Such an approach is the default download location of the alternative package, `neonUtilities`, although this can be configured there as well.  However, doing so defeats the purpose of having a persistent store, which allows us to both preserve the original raw data files between R sessions.  Because new NEON data files are released regularly, this approach allows us to easily download only the new files.  This contrasts to the typical workflow in `neonUtilities`, in which a user discards the individually downloaded files and is left to their own devices on how to keep store the resulting stacked data in order to avoid re-downloading each time.

The best way to set the file directory of the store is to declare the path using the environmental variable, `NEONSTORE_HOME`, for example, using a `.Renviron` file in the user or system home directory.  For one-off use, all functions that access the store take an optional argument `dir` to specify the directory location, (which defaults to `neon_dir()`, a function that either uses the value set by `NEONSTORE_HOME` or the `rappdirs` location if unset).  However, using the environmental variable instead results in more portable code, by allowing other users to set their own storage locations independently instead of hard-wiring it into the R code.

@@

The workhorse function of `neonstore` is `neon_download()`, which downloads all files identified by a given product code, optionally constrained to specific site, date range, or file name pattern.  For instance, we can download all of the NEON landbird survey data, identified as NEON data product `DP1.10003.001`: (The three-digit code identifies this as a Level 1 data product (DP1), product number 10003, revision 001, indicating this is the first and so far the only protocol used for bird survey sampling).

@@

## Interacting with downloaded files

Even adding a handful of products can generate thousands or hundreds of thousands of files in a local store. Surprisingly, these can even exceed the capacity of common bash commands like `ls *.csv`.  While a directory structure could mitigate this issue, `neonstore` sidesteps this issue through a convenient R-based interface.  The function `neon_index()` will list all recognized files found in the local store as a `data.frame`.  `neon_index()` uses a sophisticated filename parser, `neon_parse_filenames()`, that understands all documented file naming conventions listed at <https://data.neonscience.org/file-naming-conventions>, and quite a few that are not yet listed (internal NEON communication).  The most common of these are displayed in the table:

```{r}
neon_index()
```

`neon_index()` takes optional arguments such as `product`, `start_date`, `end_date`, and `site` to allow for convenient sub-setting.  Because most observational and instrument-based products are divided by site and month, this is still not a convenient way to browse the data.  The helper function, `neon_store()`, summarizes products that all share the same file description and can thus be easily stacked.  Like `neon_index()`, optional arguments can filter the results to specific products, dates, and sites.

```{r}
neon_store()
```

In this example, we recognize several tables corresponding to the NEON bird survey, product `DP01.10003.001` (all prefixed with `brd_`).  We can read in all the individual files
